Modelling the City of Melbourne Pedestrian Data

### 2017-05-06 presentation/proposal_pres.Rmd
initial code for imputing missing values
working draft (very much a WIP)

---
title: "Modelling the City of Melbourne Pedestrian Data"
author: "Gavin Chin, supervised by Di Cook"
date: "04/05/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readr)
library(dplyr)
library(purrr)
library(ggplot2)
library(ggmap)
library(knitr)
library(lubridate)

  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

@@

## Introduction

The City of Melbourne provides an open data platform to access council datasets through its Open Data Platform  
This project focuses primarily on the pedestrian data collected from pedestrian sensors placed throughout Melbourne's CBD

## Pedestrian Sensors

```{r, echo = F, fig.align= "center", cache=TRUE}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

## Time Series of Pedestrian Activity

```{r, message= FALSE, warning= FALSE, fig.align= "center"}
ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13") %>%
  ggplot() + geom_line(aes(x = MTime, y = Hourly_Counts)) +
      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

>>>>

---
title: "Modelling the City of Melbourne Pedestrian Data"
subtitle: "Thesis Proposal"
author: "Gavin Chin, supervised by Di Cook"
date: "8 May 2016"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, echo = FALSE, warning=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(ggmap)
library(knitr)
library(pander)
library(lubridate)
  ped_df <- read_csv("../data/Pedestrian_volume__updated_monthly_.csv")

@@

class: center

# Introduction

The City of Melbourne has an Open Data Platform which makes council data available to the public

This project will focus on the pedestrian sensor data

```{r, echo = FALSE, fig.height=5}
ggmap(melb) + geom_point(aes(x = Longitude, y = Latitude), data = ped_loc, colour = "red", alpha = 0.7) +
              labs(title = "Melbourne Pedestrian Sensor Locations") +
              xlab(NULL) + ylab(NULL) +
              theme(plot.title = element_text(hjust = 0.5), plot.caption = element_text(hjust = 0.5),
                    axis.text.x = element_blank(), axis.text.y = element_blank(),
                    axis.ticks = element_blank())
```

---

# How is the data collected?

The City of Melbourne has `r nrow(ped_loc)` sensors installed around Melbourne's CBD  
These sensors are positioned under an awning or street pole, and count the number of people passing each hour

---

# What does the data look like?

```{r}
glimpse(ped_df)
```

```{r, echo = FALSE}
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

```

---

class: center, middle

## Time series at each sensor location

```{r, message= FALSE, warning= FALSE, fig.height = 5, fig.width = 12, echo = FALSE, fig.align= "center"}
bsplot <- ped_df %>% filter(Year == "2016", Month == "June", Mdate < 25, Sensor_ID == "13")
ggplot(bsplot) + geom_line(aes(x = MTime, y = Hourly_Counts)) + ggtitle("Hourly Pedestrian Count at Flagstaff Station", subtitle = "during June 2016") +
      xlab("Hour of the Month") + ylab("Hourly Counts") + theme_light()
```

---
# Why do we want to model pedestrian traffic?

Being able to predict how many people pass through a certain location can be used by the government sector as well as the private sector

Different ways it can be used:

  Government example uses:  
    - Infrastructure planning
    - Security planning
  Private example uses:
    - Marketing campaign planning
    - Resource management
    - Investment planning

  and many other various ways..

---
class: center, inverse

# Methodology

The project has been split into multiple steps:  
  - Data Cleaning
  - Modelling
  - Visualisation
  - Shiny app development

---

# Data Cleaning

While the data variables themselves are relatively tidy, there is a large number of observations we need to impute

--

```{r, echo = FALSE, fig.align = "center", fig.height=6, fig.width=12}
ped_plot <- aggregate(x = ped_df$Hourly_Counts, by = list("Sensor_ID" = ped_df$Sensor_ID, "Year" = ped_df$Year, "Month" = ped_df$Month), FUN = sum)

ped_plot$counts <- ped_plot$x

ped_plot$Time <- parse_date(paste(ped_plot$MDate, ped_plot$Month, ped_plot$Year, sep = "-"), format = "%D-%B-%Y")

ped_plot$ID <- as.integer(ped_plot$Sensor_ID)

ped_plot$Sensor_ID <- as.factor(ped_plot$Sensor_ID)

ggplot(filter(ped_plot)) + geom_line(aes(x = Time, y = counts, colour = Sensor_ID)) + theme_light() + theme(legend.position = "none") + labs(x = "Year", y = "Pedestrian Counts") + ggtitle("Monthly Total Counts by Sensor")

```

---

# R Code

```{r comment='#'}
# a boring regression
fit = lm(dist ~ 1 + speed, data = cars)
coef(summary(fit))
dojutsu = c('地爆天星', '天照', '加具土命', '神威', '須佐能乎', '無限月読')
grep('天', dojutsu, value = TRUE)
```

---

# R Plots

```{r cars, fig.height=4, dev='svg'}
par(mar = c(4, 4, 1, .1))
plot(cars, pch = 19, col = 'darkgray', las = 1)
abline(fit, lwd = 2)
```

---

# HTML Widgets

I have not thoroughly tested HTML widgets against **xaringan**. Some may work well, and some may not. It is a little tricky.

Similarly, the Shiny mode (`runtime: shiny`) does not work. I might get these issues fixed in the future, but these are not of high priority to me. I never turn my presentation into a Shiny app. When I need to demonstrate more complicated examples, I just launch them separately. It is convenient to share slides with other people when they are plain HTML/JS applications.

See the next page for two HTML widgets.

---

```{r out.width='100%', fig.height=6, eval=require('leaflet')}
library(leaflet)
leaflet() %>% addTiles() %>% setView(mean(ped_loc$Longitude), mean(ped_loc$Latitude), zoom = 14)
```

---

```{r eval=require('DT'), tidy=FALSE}
DT::datatable(
  head(iris, 10),
  fillContainer = FALSE, options = list(pageLength = 5)
)
```

---

# Some Tips

- When you use the "Infinite Moon Reader" addin in RStudio, your R session will be blocked by default. You can click the red button on the right of the console to stop serving the slides, or use the _daemonized_ mode so that it does not block your R session. To do the latter, you can set the option

    ```r
    options(servr.daemon = TRUE)
    ```

    in your current R session, or in `~/.Rprofile` so that it is applied to all future R sessions. I do the latter by myself.

    To know more about the web server, see the [**servr**](https://github.com/yihui/servr) package.

--

- Do not forget to try the `yolo` option of `xaringan::moon_reader`.

    ```yaml
    output:
      xaringan::moon_reader:
        yolo: true
    ```

---

# Some Tips

- Slides can be automatically played if you set the `autoplay` option under `nature`, e.g. go to the next slide every 30 seconds (this can be useful for lightning talks):

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          autoplay: 30000
    ```

--

- The option `highlightLines: true` of `nature` will highlight code lines that start with `*` or are wrapped in `{{ }}`;

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          highlightLines: true
    ```

    See examples on the next page.

---

# Some Tips

.pull-left[
An example using a leading `*`:

    ```r
    if (TRUE) {
    ** message("Very important!")
    }
    ```
Output:
```r
if (TRUE) {
* message("Very important!")
}
```

This is invalid R code, so it is a plain fenced code block that is not executed.
]

.pull-right[
An example using `{{}}`:

    `r ''````{r tidy=FALSE}
    if (TRUE) {
    *{{ message("Very important!") }}
    }
    ```
Output:
```{r tidy=FALSE}
if (TRUE) {
{{ message("Very important!") }}
}
```

It is valid R code so you can run it.
]

---

# Some Tips

- To make slides work offline, you need to download a copy of remark.js in advance, because **xaringan** uses the online version by default (see the help page `?xaringan::moon_reader`).

- You can use `xaringan::summon_remark()` to download the latest or a specified version of remark.js. By default, it is downloaded to `libs/remark-latest.min.js`.

- Then change the `chakra` option in YAML to point to this file, e.g.

    ```yaml
    output:
      xaringan::moon_reader:
        chakra: libs/remark-latest.min.js
    ```

- If you used Google fonts in slides (the default theme uses _Yanone Kaffeesatz_, _Droid Serif_, and _Source Code Pro_), they won't work offline unless you download or install them locally. The Heroku app [google-webfonts-helper](https://google-webfonts-helper.herokuapp.com/fonts) can help you download fonts and generate the necessary CSS.

---

# CSS

Among all options in `xaringan::moon_reader`, the most challenging but perhaps also the most rewarding one is `css`, because it allows you to customize the appearance of your slides using any CSS rules or hacks you know.

You can see the default CSS file [here](https://github.com/yihui/xaringan/blob/master/inst/rmarkdown/templates/xaringan/resources/example.css). You can completely replace it with your own CSS files, or define new rules to override the default.

---

# CSS

For example, suppose you want to change the font for code from the default "Source Code Pro" to "Ubuntu Mono". You can create a CSS file named, say, `ubuntu-mono.css`:

```css
@import url(https://fonts.googleapis.com/css?family=Ubuntu+Mono:400,700,400italic);

.remark-code, .remark-inline-code { font-family: 'Ubuntu Mono'; }
```

Then set the `css` option in the YAML metadata:

```yaml
output:
  xaringan::moon_reader:
    css: ["default", "ubuntu-mono.css"]
```

Here I assume `ubuntu-mono.css` is under the same directory as your Rmd.

---
class: inverse, middle, center
background-image: url(https://upload.wikimedia.org/wikipedia/commons/3/39/Naruto_Shiki_Fujin.svg)
background-size: contain

# Naruto

---
background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)
background-size: 100px
background-position: 90% 8%

# Sharingan

The R package name **xaringan** was derived<sup>1</sup> from **Sharingan**, a dōjutsu in the Japanese anime _Naruto_ with two abilities:

- the "Eye of Insight"

- the "Eye of Hypnotism"

I think a presentation is basically a way to communicate insights to the audience, and a great presentation may even "hypnotize" the audience.<sup>2,3</sup>

.footnote[
[1] In Chinese, the pronounciation of _X_ is _Sh_ /ʃ/ (as in _shrimp_). Now you should have a better idea of how to pronounce my last name _Xie_.

[2] By comparison, bad presentations only put the audience to sleep.

[3] Personally I find that setting background images for slides is a killer feature of remark.js. It is an effective way to bring visual impact into your presentations.
]

---

# Naruto terminology

The **xaringan** package borrowed a few terms from Naruto, such as

- [Sharingan](http://naruto.wikia.com/wiki/Sharingan) (写輪眼; the package name)

- The [moon reader](http://naruto.wikia.com/wiki/Moon_Reader) (月読; an attractive R Markdown output format)

- [Chakra](http://naruto.wikia.com/wiki/Chakra) (查克拉; the path to the remark.js library, which is the power to drive the presentation)

- [Nature transformation](http://naruto.wikia.com/wiki/Nature_Transformation) (性質変化; transform the chakra by setting different options)

- The [infinite moon reader](http://naruto.wikia.com/wiki/Infinite_Tsukuyomi) (無限月読; start a local web server to continuously serve your slides)

- The [summoning technique](http://naruto.wikia.com/wiki/Summoning_Technique) (download remark.js from the web)

You can click the links to know more about them if you want. The jutsu "Moon Reader" may seem a little evil, but that does not mean your slides are evil.

---

class: center

# Hand seals (印)

Press `h` or `?` to see the possible ninjutsu you can use in remark.js.

![](https://upload.wikimedia.org/wikipedia/commons/7/7e/Mudra-Naruto-KageBunshin.svg)

---

class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
